### 2022-01-18 src-latex/main-paper.tex
WIP: section 4

\section*{Thanks}
Laureencia for annotations, Guillaume Thomas for data processing?

\section*{Acknowledment}
This work is supported by the French National Research Agency (ANR), as part of the SODUCO project, under Grant ANR-18-CE38-0013.

>>>>

\section*{Acknowledments and Thanks}
This work is supported by the French National Research Agency (ANR), as part of the SODUCO project, under Grant ANR-18-CE38-0013.

The authors want to thank \textcolor{green}{Laureencia for annotations, Guillaume Thomas for data processing? Stéphane Bacciochi and Pascal Cristofoli for helping to create the gold reference dataset ?}

### 2022-01-18 src-latex/parts/30-dataset.tex
WIP: section 4

To build our ground truth corpus, we annotated a few pages of each of the major directory series for which we have digitised copies, using a tool developed for our project (see Figure \ref{fig:annotator}). First, the image of a given directory page is loaded. Second, entries are automatically detected and boxes are drawn around them as shown in the left panel of the Graphical User Interface (GUI). Then the text of each entry is extracted by an OCR tool and finally named entities are annotated by a NER tool: the results are displayed respectively in the first and second tabs of the right panel of the GUI. The entry detection, OCR and NER tasks are performed automatically. But as they are prone to errors, their results can be corrected manually through the GUI. In the end, the entry bounding boxes, their text, and the named entities are stored in a JSON file.

@@

\begin{itemize}
    \item statistics about entry lengths (histogram of number of char / tokens par entry)
    \item entities we considered and how we annotated them (how we assessed gold's quality)
    \item statistics on the distribution of entities across directories and their arrangement within entries
    \item OCR data (automated systems and clean text + how we created the "clean" text + whether/how we assessed clean text's quality)
    \item Measure inter-annotator agreement ? 
    \item Explain charset projection
    \item \dots
\end{itemize}

>>>>

To build our ground truth corpus, we annotate a few pages of each of the major directory series for which we have digitised copies, using a tool developed for our project (see Figure \ref{fig:annotator}). First, the image of a given directory page is loaded. Second, entries are automatically detected and boxes are drawn around them as shown in the left panel of the Graphical User Interface (GUI). Then the text of each entry is extracted by an OCR tool and finally named entities are annotated by a NER tool: the results are displayed respectively in the first and second tabs of the right panel of the GUI. The entry detection, OCR and NER tasks are performed automatically. But as they are prone to errors, their results can be corrected manually through the GUI. In the end, the entry bounding boxes, their text, and the named entities are stored in a JSON file.

@@

\begin{itemize}
    \item statistics about entry lengths (histogram of number of char / tokens par entry)
    \item entities we consider and how we annotated them (how we assessed gold's quality)
    \item statistics on the distribution of entities across directories and their arrangement within entries
    \item OCR data (automated systems and clean text + how we created the "clean" text + whether/how we assess clean text's quality)
    \item Measure inter-annotator agreement ? 
    \item Explain charset projection
    \item \dots
\end{itemize}

### 2022-01-18 src-latex/parts/40-ner-approaches.tex
Reorganised end of section 5 to save space
WIP: section 4

% !TeX root = ../main-paper.tex
\section{NER Approaches}

In our tests, we have selected two deep learning-based NER models trained on French texts: Spacy 3 NER pipeline and CamemBERT. They are adapted to  historical texts as follows.

\subsubsection{Spacy 3 NER pipeline}
The \textit{of the shelf} NLP library SpaCY offers a named entity recognition tool based on an ad hoc architecture which has not been yet published but is explained by the developers on their website. Although the latest version of SpaCy (v3) leverages BERT models, we could not use them as no NER pipeline is available for the French language at the time of our experiments\footnote{https://spacy.io/models/fr}.
The vanilla NER pipeline is two-fold. Words are first encoded into local context-aware embeddings using a window-based CNN similar to~\cite{collobert2011}.
The decision layer is an adaptation of the transition-based model presented in~\cite{lample2016}.
As words are processed sequentially, their vectors are concatenated with those of the last known entities to encode the nearby predicted semantics as a dynamic attention mechanism.
The classification layer relies on a finite-state machine whose transition probabilities are learned using a multilayer perceptron.
In 2018~\cite{won2018} compared SpaCy's vanilla NER performance on historical texts about places with several other software packages (Stanford-NER, Ner-Tagger, Edinburgh Geopoarser and Polyglot). Results showed SpaCy to be average with a F1-score of 0.57.
The SpaCy developers claimed an accuracy of 0.85 for the English NER pipeline \textit{en\_core\_web\_lg} on the OntoNotes 5.0 corpus\footnote{https://spacy.io/usage/facts-figures}.

\subsubsection{CamemBERT \& CamemBERT pretrained on raw entries}

Language model embeddings pretrained using Transformers, like BERT, have become a new paradigm for NER\cite{li2020}. Indeed, these embeddings can be used as distributed representations for input instead of traditional embeddings like Google Word2Vec and they can be further fine-tuned for NER task by adding an additional output layer. Moreover, they can also be further pretrained in an unsupervised way on historical texts for domain adaptation.

As our directories are written in French, we chose the language model CamemBERT \cite{martin-etal-2020-camembert}, a Transformer model trained on a French corpus. We started from the CamemBERT-ner model already fine-tuned on WikiNER-fr corpus and published on the Hugging Face repository \footnote{\url{https://huggingface.co/Jean-Baptiste/camembert-ner}}. Its output layer is a linear model with a Softmax function.

>>>>

% !TeX root = ../main-paper.tex
\section{NER Approaches}
We select twe deep-learning based NER models available in packaged software libraries: SpaCy NLP pipelines and CamemBERT.
In the case of CamemBERT, we also create an additional model pre-trained on a collection of unannotated directories entries extracted with Pero-OCR.
The three models are then fine-tuned for named entity recognition in historical directories following a procedure detailed in sections~\ref{subsection:experiment-1} and ~\ref{subsection:experiment-2}.

\subsubsection{SpaCy NER pipeline}
Spacy is a software library that offers NLP components assembled in modular pipelines specialised by language.
Although BERT is available in the latest version of SpaCy (v3), the pipeline for French does not provide a NER layer.
Hence we rely on SpaCy's ad hoc pipelines trained on French corpora and capable of named entity recognition.
% deep-sequoia and wikiner-fr, capable of named entity recognition.\textit{fr\_core\_news\_lg}\footnote{https://spacy.io/models/fr} trained two corpora in French: deep-sequoia and wikiner-fr, capable of named entity recognition.
The global architecture of these pipelines have not been yet published but is explained by the developers on their website.
Words are first encoded into local context-aware embeddings using a window-based CNN similar to~\cite{collobert2011}.
The decision layer is an adaptation of the transition-based model presented in~\cite{lample2016}.
As words are processed sequentially, their vectors are concatenated with those of the last known entities to encode the nearby predicted semantics.
The classification layer relies on a finite-state machine whose transition probabilities are learned using a multilayer perceptron.
\bertrand{Drop the next sentences if we need space}
In 2018~\cite{won2018} evaluated SpaCy's NER ability to detect place names in five corpora of ancient letters written in English.
They measured an average F1 score of 0.57.
The SpaCy developers claim an accuracy of 0.85 for the English NER pipeline \textit{en\_core\_web\_lg} on the OntoNotes 5.0 corpus\footnote{https://spacy.io/usage/facts-figures}.
%For our experiments, the French model \textit{fr\_core\_news\_lg} is fine-tuned using our ground truth corpus.

\subsubsection{CamemBERT}
Language model based on Transformers, like BERT, have become a new paradigm for NER\cite{li2020}. 
The learned embeddings can be used as distributed representations for input instead of traditional embeddings like Google Word2Vec, and they can be further fine-tuned for NER by adding an additional output layer. 
They can also be pretrained in an unsupervised way on historical texts for domain adaptation.

As our directories are written in French, we chose the language model CamemBERT \cite{martin-etal-2020-camembert}, a Transformer model trained on a French corpus.
\bertrand{WIP}

### 2022-01-18 src-latex/parts/50-xp-setup.tex
WIP: section 4

\subsection{NER sensibility to the number of training samples}
The first experiment evaluates the performances of the three models on training sets of different sizes.
To do so, we split the gold reference into a training set, a development set, and a test set. The training set is then gradually reduced in size while maintaining the relative frequency of directories within.
The training and testing procedure is the same for the three models.

@@

\subsection{NER in noisy OCRed texts}
The second experiment aims to evaluate the impact of the OCR used on the quality of the NER results. To this end, we use directory entries whose text has been extracted, corrected and annotated manually (see Section 3). This reference text is aligned with the text produced on the same entries by Tesseract and PERO OCR. The named entities annotated in the reference text are then projected into the OCRed texts. To make sure that all training, development and test datasets contain the same annotations, only the entries for which the same number of annotations are found in the reference text and in the OCRised texts are kept. Finally, the reference text is divided into three datasets, using the same method as in Experiment 1. Two additional test files are generated with exactly the same entries as the test file produced from the reference text: one from the text produced by PERO OCR and one from that produced by Tesseract.

@@

\subsubsection{OCR systems parameters}
\begin{itemize}
    \item original OCR
    \item Tesseract 4
    \item Pero
    \item other?
\end{itemize}

\subsubsection{Spacy}
We use Spacy\mcite{spacy} for the CNN implementation
and Huggingface for the BERT 
tok2vec(words embeddings + encoding) + attention layer  +  transition-based model.

\subsubsection{Huggingface CamemBERT}
\paragraph{Huggingface CamemBERT}
CamemBERT + linear classifier as decision layer

EX1: Pour chaque modèle train sur les 8 datasets * 5 runs, avec mêmes paramètres pour l’entraînement: early stopping  (patience = 3)

Ex 2: On garde CamemBERT et CamemBERT pretrained fine-tuned sur le train clean (patience = 5), validé avec dev clean. Teste les deux modèles sur les trois datasets de test.

>>>>

\subsection{NER sensibility to the number of training samples}
\label{subsection:experiment-1}
The first experiment evaluates the performances of the three models on training sets of different sizes.
To do so, we split the gold reference into a training set, a development set, and a test set. The training set is then gradually reduced in size while maintaining the relative frequency of directories within.
The training and testing procedure is the same for the three models.

@@

\subsection{NER in noisy OCRed texts}
\label{subsection:experiment-2}
The second experiment assesses the impact of the OCR used on the quality of the NER results. To this end, we use directory entries whose text has been extracted, corrected and annotated manually (see Section 3). This reference text is aligned with the text produced on the same entries by Tesseract and PERO OCR. The named entities annotated in the reference text are then projected into the OCRed texts. To make sure that all training, development and test datasets contain the same annotations, only the entries for which the same number of annotations are found in the reference text and in the OCRised texts are kept. Finally, the reference text is divided into three datasets, using the same method as in Experiment 1. Two additional test files are generated with exactly the same entries as the test file produced from the reference text: one from the text produced by PERO OCR and one from that produced by Tesseract.

@@

\subsubsection{OCR systems parameters}
\nathalie{Y a-t-il des précisions à ajouter sur la façon dont ont été utilisés les OCR? Ou bien ce sont les versions "sur étagère" qui ont servi?}
\begin{itemize}
    \item Tesseract 4
    \item Pero
\end{itemize}

\subsubsection{Spacy}
We used the French model \textit{fr\_core\_news\_lg} provided by Spacy\mcite{spacy} library and leveraged the library utilities to fine-tune it with various training and validation datasets derived from our ground truth dataset (see section 5.2 for details). The base model is fine-tuned in turn with each of the 8 generated training data sets, using the same training parameters. Early stopping is activated with patience parameters set at 1600. For each size of training dataset, 5 runs are performed, always with the same parameters, and the mean value of each performance score is computed.

%and Huggingface for the BERT 
%tok2vec(words embeddings + encoding) + attention layer  +  transition-based %model.

\subsubsection{Huggingface CamemBERT}
We started from the CamemBERT-ner model already fine-tuned on WikiNER-fr corpus and published on the Hugging Face repository \footnote{\url{https://huggingface.co/Jean-Baptiste/camembert-ner}}. Its output layer is a linear model with a Softmax function. Two NER models are derived from this base model.

\textcolor{green}{Ex 2: On garde CamemBERT et CamemBERT pretrained fine-tuned sur le train clean (patience = 5), validé avec dev clean. Teste les deux modèles sur les trois datasets de test.}

The base model is pretrained on 845 000 entries extracted with PERO OCR for the Next  Sentence  Prediction  (NSP) and  Masked  Language  Model(MLM) tasks. Then it is fine-tuned with the same parameters and on the same training datasets as the first model.

The source code of all our experiments is publicly available on the following repository \url{lien github}.
